---
title: "PR-758: H2 effects on CH4 lifetime"
date: "`r Sys.Date()`"
output:
  github_document:
    toc: true
    toc_depth: 2
---

Generates the plots for https://github.com/JGCRI/hector/pull/758. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```


```{r, warning=FALSE, message=FALSE}
BASE_DIR <- here::here("dev", "PR-758")

library(dplyr)

#HECTOR_DIR <- "/Users/dorh012/Documents/2024/H2Materials/Hector_H2_Project/../hector"
#devtools::load_all(HECTOR_DIR) # this is a specific version of Hector
#TODO replace with the commit ssha 
remotes::install_github("jgcri/hector@ch4_oh_lifetime_only")
library(ggplot2)
theme_set(theme_bw())
```

## Functions

Demonstrate that H2 related functions functions work as expected. 

```{r}
getunits(EMISSIONS_H2())

# The control run
ini_file <- file.path(BASE_DIR, "inputs", "picontrol_ch4-emiss.ini")
hc <- newcore(ini_file, name = "control")
run(hc)
fetchvars(hc, NA, COEFF_H2())
```



## Different H2 Impulse Sizes 

Demonstrate how changing the size of the H2 emissions pulse effects Hector results throughout the climate system. Since we have only implemented H2 impacts on CH4 lifetime these results are going to be somewhat small. 

```{r}
# Information that is constant for for all of our hector runs
dates <- 1750:2000
vars <- c(EMISSIONS_CH4(), EMISSIONS_H2(), LIFETIME_OH(), CONCENTRATIONS_CH4(),
          RF_CH4(), RF_H2O_STRAT(), RF_O3_TROP(), RF_TOTAL(), GLOBAL_TAS())
ini_file <- file.path(BASE_DIR, "inputs", "picontrol_ch4-emiss.ini")
IMPULSE_YR <- 1850
```



```{r}
# The control run
hc <- newcore(ini_file, name = "control")
run(hc)
out1 <- fetchvars(hc, dates, vars)
```

```{r}
# Impulse 1
H2_PULSE <- 200
hc <- newcore(ini_file, name = paste0("h2 impulse: ", H2_PULSE))
setvar(hc, IMPULSE_YR, var =  EMISSIONS_H2(), values =H2_PULSE,
       unit = getunits(EMISSIONS_H2()))
reset(hc)
run(hc)
out2 <- fetchvars(hc, dates, vars)
```

```{r}
# Impulse 2
H2_PULSE2 <- H2_PULSE * 2
hc <- newcore(ini_file, name = paste0("h2 impulse: ", H2_PULSE2))
setvar(hc, IMPULSE_YR, var =  EMISSIONS_H2(), values = H2_PULSE2,
       unit = getunits(EMISSIONS_H2()))
reset(hc)
run(hc)
out3 <- fetchvars(hc, dates, vars)
```


```{r}
# Combine all of the results into a single data frame
rbind(out1, out2, out3) %>%
    mutate(year = year - IMPULSE_YR) %>%
    filter(year >= -2) %>%
    filter(year <= 50) ->
    to_plot

# Make sure that the plot is in the correct order
to_plot$variable <- factor(to_plot$variable, levels = vars)

to_plot %>%
    ggplot(aes(year, value, color = scenario)) +
    geom_line() +
    facet_wrap("variable", scales = "free") +
    labs(x = NULL, y = NULL) +
    theme(legend.title = element_blank())
```

```{r, echo=FALSE}
# remove the results dfs 
remove("out1", "out2", "out3", "to_plot")
```


## Different H2 Coefficent 

Demonstrate how changing the H2 coefficient can change tau OH and [CH4]

```{r}
# Information that is constant for for all of our hector runs
dates <- 1750:2000
vars <- c(EMISSIONS_CH4(), EMISSIONS_H2(), LIFETIME_OH(), CONCENTRATIONS_CH4(),
          RF_CH4(), RF_H2O_STRAT(), RF_O3_TROP(), RF_TOTAL(), GLOBAL_TAS())
ini_file <- file.path(BASE_DIR, "inputs", "picontrol_ch4-emiss.ini")
IMPULSE_YR <- 1850
```

Default H2 coefficient 

```{r}
# Impulse 1, default H2 coeff 
H2_PULSE <- 200
hc <- newcore(ini_file)
setvar(hc, IMPULSE_YR, var =  EMISSIONS_H2(), values = H2_PULSE, unit = getunits(EMISSIONS_H2()))
reset(hc)
run(hc)
out1 <- fetchvars(hc, dates, vars)
default_ch2 <- fetchvars(hc, NA, COEFF_H2())[["value"]]
out1$scenario <- "defualt coef H2"
out1$coef <- default_ch2
```

Lower H2 coefficient 

```{r}
setvar(hc, NA, COEFF_H2(), values = default_ch2 * 0.5, unit = "(undefined)")
reset(hc)
run(hc)
out2 <- fetchvars(hc, dates, vars)
new_ch2 <- fetchvars(hc, NA, COEFF_H2())[["value"]]
out2$scenario <- "lower coef H2"
out2$coef <- new_ch2
```


Higher H2 coefficient 

```{r}
setvar(hc, NA, COEFF_H2(), values = default_ch2 * 1.5, unit = "(undefined)")
reset(hc)
run(hc)
out3 <- fetchvars(hc, dates, vars)
new_ch2 <- fetchvars(hc, NA, COEFF_H2())[["value"]]
out3$scenario <- "higher coef H2"
out3$coef <- new_ch2
```

```{r}
# Format output for plotting
rbind(out1, out2, out3) %>%  
    mutate(year = year - IMPULSE_YR) %>% 
    filter(year >= -2 & year <= 40) ->
     to_plot
```

Confirm that the emissions are correct 

```{r}
to_plot %>% 
    filter(variable == EMISSIONS_H2()) %>%  
    ggplot(aes(year, value, color = scenario, linetype = scenario)) + 
    geom_line(size = 1) + 
    labs(y = getunits(EMISSIONS_H2()), x = NULL) + 
    theme(legend.title = element_blank())
```


```{r}
to_plot %>% 
    filter(variable %in% c(LIFETIME_OH(), CONCENTRATIONS_CH4(), GLOBAL_TAS())) -> 
    to_plot

plot_these_vars <- c(LIFETIME_OH(), CONCENTRATIONS_CH4(), GLOBAL_TAS())
to_plot$variable <- factor(to_plot$variable , levels = plot_these_vars)

to_plot %>% 
    ggplot(aes(year, value, color = scenario, linetype = scenario)) + 
    geom_line(size = 1) + 
    labs(y = NULL, x = NULL) + 
    facet_wrap("variable", scales = "free", ncol = 1) + 
    theme(legend.title = element_blank())



```




```{r, echo=FALSE}
# remove the results dfs 
remove("out1", "out2", "out3", "to_plot")
```


## GWP100 

Sand et al. report the AWGP 100 for H2 from various ESMs have included results broken down by each of the three indirect $H_2$ effects so we can use that to compare our $H_2$ implementation here. 

$$\text{GWP}_{100} = \frac{\int_{0}^t \Delta \text{RF H}_2}{\int_{0}^t \Delta \text{RF CO}_2}$$
where the $\Delta \text{RF H}_2$ is from a pulse of 1 Tg of $\text{H}_2$ and $\Delta \text{RF CO}_2}$ is from a pulse of 1 Tg of $\text{CO}_2}$. 

### CO2 pulse runs

Since we don't have a emission driven pi control (for CO2) we have to apply the pulse on top of a control run. 

```{r}
# Set information we would like to collect from the hector runs
dates <- 1750:2100
vars <- c(c(RF_CO2(), RF_TOTAL(), GLOBAL_TAS(), LIFETIME_OH(), RF_O3_TROP(),
            FFI_EMISSIONS(), RF_H2O_STRAT()))

# Our control run 
inifile <- file.path(HECTOR_DIR, "inst", "input", "hector_ssp245.ini")
hc <- newcore(inifile, name = "control")
run(hc)
fetchvars(hc, dates, vars) %>%
    rename(control = value) ->
    out1
shutdown(hc)

# Impulse run
IMPULSE_YR <- 1850
inifile <- file.path(HECTOR_DIR, "inst", "input", "hector_ssp245.ini")
hc <- newcore(inifile, name = "impulse")
run(hc)
contorl_emiss_val <- fetchvars(hc, IMPULSE_YR, FFI_EMISSIONS())$value
# We want our pulse to only equivalent to 1 Tg of CO2 but Hector inputs are in 
# Pg C so need to convert before adding the pulse to 
CO2_Tg <- 1 
CONVERSION_FACTOR <- (12.01/44.01) * 1e-3 # convert from Tg CO2 to PgC 
impulse <- contorl_emiss_val + (CO2_Tg * CONVERSION_FACTOR)
setvar(hc, IMPULSE_YR, FFI_EMISSIONS(), impulse, getunits(FFI_EMISSIONS()))
reset(hc)
run(hc)
fetchvars(hc, dates, vars) %>%
    rename(impulse = value) ->
    out2
shutdown(hc)

# Becasue of the way that Hector's carbon cycle runs offset the impulse year by 1 
out1 %>%
    cbind(impulse = out2$impulse) %>%
    mutate(value = impulse - control) %>%
    mutate(year = year - (IMPULSE_YR+1)) %>%
    filter(year >= 0) %>%
    filter(year <= 100) %>%
    select(year, variable, value, units) ->
    co2_100_rslts
```


### H2 pulse runs

```{r}
# Information that is constant for for all of our hector runs
vars <- c(EMISSIONS_H2(), EMISSIONS_CH4(), CONCENTRATIONS_CH4(), LIFETIME_OH(),
          RF_CH4(), RF_TOTAL(), GLOBAL_TAS(),
          RF_H2O_STRAT(), RF_O3_TROP(), RF_CO2())
ini_file <- file.path(BASE_DIR, "inputs", "picontrol_ch4-emiss.ini")
IMPULSE_YR <- 1850

# Impulse 1
H2_PULSE <- 1
hc <- newcore(ini_file, name = paste0("h2 impulse: ", H2_PULSE))
setvar(hc, IMPULSE_YR, var =  EMISSIONS_H2(), values =H2_PULSE,
       unit = getunits(EMISSIONS_H2()))
reset(hc)
run(hc)

# Format output 
fetchvars(hc, dates, vars) %>%
    mutate(year = year - IMPULSE_YR) %>%
    filter(year >= 0) %>%
    filter(year <= 100) %>%
    select(year, variable, value, units) ->
    h2_100_rslts
```

Integrate the pulse run results and calculate the AGWP100 

```{r}
# Calculate the GWP100 as the ratio of the cumulative RF change 100 years (integrated)
co2_integral <- summarise(co2_100_rslts, value = sum(value), .by = c("variable"))
h2_integral <- summarise(h2_100_rslts, value = sum(value), .by = c("variable"))

# This is consistent with the model mean GWP 100 due to CH4 from Sand et al.
# Supplementary Table 6
GWP100 <- round(h2_integral$value[h2_integral$variable == "RF_tot"] / co2_integral$value[co2_integral$variable == "RF_tot"], 3)


paste0("GWP100 H2: ", GWP100)
```


Looking at Sand et al. Supplementary Table 6 the GWP100 of hydrogen due to methane ranges from 4.9 to 5.7 with a model mean of 5.1, which is not too far off what we are getting from Hector at the moment! 


RF response to the H2 pulse

```{r}
h2_100_rslts$type <- "H2"
co2_100_rslts$type <- "CO2"


h2_100_rslts %>%
    bind_rows(co2_100_rslts) %>%
    filter(variable == RF_TOTAL()) %>%
    ggplot(aes(year, value, color = type))+
    geom_line() +
    labs(title = "RF from 1Tg Emissions",
         subtitle = paste0("GWP100: ", GWP100),
         caption = "H2 impacts Tau OH",
         x = "years since emissions pulse",
         y = "Total RF (Wm-2)") +
    theme(legend.title = element_blank())
```
