---
title: "PR-775: H2 effects on Strat H2O Vapor RF"
date: "`r Sys.Date()`"
output:
  github_document:
---

Make plots for https://github.com/JGCRI/hector/pull/775 where we add the H2 effects on strat water vapor radiative forcing. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, 
                       out.width='80%', fig.align='center', class='center', 
                      fig.path = "figs/")
```


```{r, warning=FALSE, message=FALSE, warning=FALSE}
BASE_DIR <- here::here("dev", "PR-775")

library(dplyr)

# TODO this needs to be a specific version of Hector will want to replace with the 
# git ssha after the PR is approved
HECTOR_DIR <- "/Users/dorh012/Documents/2024/H2Materials/Hector_H2_Project/../hector"
devtools::load_all(HECTOR_DIR) # this is a specific version of Hector
library(ggplot2)
theme_set(theme_bw())
```


# Implementation

(S11) from Dorheim et al. 2024 for our previous $RF_{strat_{H_2O}}$ equation, we had 

$$ RF_{strat_{H_2O}} = 0.0485 \frac{[CH_4](t) - [CH_4]_0}{1831-[CH_4]_0}$$
which only accounts strat $H_2O$ created from methane oxidation with OH radicals. However as part of the indirect climate effects of $H_2$ we also have $H_2$ emissions contributing to $RF_{strat_{H_2O}}$. The new implementation of $RF_{strat_{H_2O}}$ is now 


$$ RF_{strat_{H_2O}} = 0.0485 \frac{[CH_4](t) - [CH_4]_0}{1831-[CH_4]_0} +  0.00019 (E_{H_2}(t) - {E_{H_2}}_0)$$
Where the coefficient of $0.00019$ W m −2 per Tg yr− 1 is the multi-model mean from Sand et al. 2023. 



# Strat H2O Vapor Effects Only 

To start let's isolate the RF strat h2o vapor effects, by making sure the H2 effects 
on CH4 are set to 0.

```{r}

# Add in this helper function for the moment since the R bindings had been depreciated
COEFF_H2 <- function(){"CH2"}


# Helper function that
my_newcore <- function(ini, name){
    hc <- newcore(ini_file, name = "control")
    setvar(hc, NA, COEFF_H2(), values = 0, unit = "(undefined)")
    return(hc)
    
}
```



```{r}
# Information that is constant for for all of our hector runs
dates <- 1750:2000
vars <- c(EMISSIONS_CH4(), EMISSIONS_H2(), CH4_LIFETIME_OH(), CONCENTRATIONS_CH4(),
          RF_CH4(), RF_H2O_STRAT(), RF_O3_TROP(), RF_TOTAL(), GLOBAL_TAS())
ini_file <- file.path(BASE_DIR, "inputs", "picontrol_ch4-emiss.ini")
IMPULSE_YR <- 1850

# The control run
hc <- newcore(ini_file, name = "control")
setvar(hc, NA, COEFF_H2(), values = 0, unit = "(undefined)")
run(hc)
out1 <- fetchvars(hc, dates, vars)


# Impulse 1
H2_PULSE <- 500
hc <- newcore(ini_file, name = paste0("h2 impulse: ", H2_PULSE))
setvar(hc, NA, COEFF_H2(), values = 0, unit = "(undefined)")
setvar(hc, IMPULSE_YR, var =  EMISSIONS_H2(), values =H2_PULSE,
       unit = getunits(EMISSIONS_H2()))
reset(hc)
run(hc)
out2 <- fetchvars(hc, dates, vars)
```
Let's take a look at the difference results from an impulse of H2 emissions in 1850 and a pi control run.  

```{r}
out2 %>% 
    mutate(value = value - out1$value) %>% 
    filter(year > 1840 & year <= 1900) %>% 
    ggplot(aes(year, value)) + 
    geom_line() + facet_wrap("variable", scales = "free") + 
    labs(title = "H2 Indirect Effects on Strat H2O Vapor RF", 
         subtitle = "No CH4 Indirect Effects", 
         y = "H2 Impulse - Pi Control", 
         x = NULL)
```




```{r, echo=FALSE}
# remove the results dfs 
remove("out1", "out2", "out3", "to_plot")
```


## GWP100 

Sand et al. report the AWGP 100 for H2 from various ESMs have included results broken down by each of the three indirect $H_2$ effects so we can use that to compare our $H_2$ implementation here. 

$$\text{GWP}_{100} = \frac{\int_{0}^t \Delta \text{RF H}_2}{\int_{0}^t \Delta \text{RF CO}_2}$$
where the $\Delta \text{RF H}_2$ is from a pulse of 1 Tg of $\text{H}_2$ and $\Delta \text{RF CO}_2}$ is from a pulse of 1 Tg of $\text{CO}_2}$. 

### CO2 pulse runs

Since we don't have a emission driven pi control (for CO2) we have to apply the pulse on top of a control run. 

```{r}
# Set information we would like to collect from the hector runs
dates <- 1750:2100
vars <- c(c(RF_CO2(), RF_TOTAL(), GLOBAL_TAS(), CH4_LIFETIME_OH(), RF_O3_TROP(),
            FFI_EMISSIONS(), RF_H2O_STRAT()))

# Our control run 
inifile <- file.path(HECTOR_DIR, "inst", "input", "hector_ssp245.ini")
hc <- newcore(inifile, name = "control")
run(hc)
fetchvars(hc, dates, vars) %>%
    rename(control = value) ->
    out1
shutdown(hc)

# Impulse run
IMPULSE_YR <- 1850
inifile <- file.path(HECTOR_DIR, "inst", "input", "hector_ssp245.ini")
hc <- newcore(inifile, name = "impulse")
run(hc)
contorl_emiss_val <- fetchvars(hc, IMPULSE_YR, FFI_EMISSIONS())$value
# We want our pulse to only equivalent to 1 Tg of CO2 but Hector inputs are in 
# Pg C so need to convert before adding the pulse to 
CO2_Tg <- 1 
CONVERSION_FACTOR <- (12.01/44.01) * 1e-3 # convert from Tg CO2 to PgC 
impulse <- contorl_emiss_val + (CO2_Tg * CONVERSION_FACTOR)
setvar(hc, IMPULSE_YR, FFI_EMISSIONS(), impulse, getunits(FFI_EMISSIONS()))
reset(hc)
run(hc)
fetchvars(hc, dates, vars) %>%
    rename(impulse = value) ->
    out2
shutdown(hc)

# Because of the way that Hector's carbon cycle runs offset the impulse year by 1 
out1 %>%
    cbind(impulse = out2$impulse) %>%
    mutate(value = impulse - control) %>%
    mutate(year = year - (IMPULSE_YR+1)) %>%
    filter(year >= 0) %>%
    filter(year <= 100) %>%
    select(year, variable, value, units) ->
    co2_100_rslts
```





### H2 pulse runs

```{r}
# Information that is constant for for all of our hector runs
vars <- c(EMISSIONS_H2(), EMISSIONS_CH4(), CONCENTRATIONS_CH4(),
          CH4_LIFETIME_OH(), RF_CH4(), RF_TOTAL(), GLOBAL_TAS(),
          RF_H2O_STRAT(), RF_O3_TROP(), RF_CO2())
ini_file <- file.path(BASE_DIR, "inputs", "picontrol_ch4-emiss.ini")
IMPULSE_YR <- 1850

# Impulse 1
H2_PULSE <- 1
hc <- my_newcore(ini = ini_file, name = paste0("h2 impulse: ", H2_PULSE))
setvar(hc, IMPULSE_YR, var =  EMISSIONS_H2(), values =H2_PULSE,
       unit = getunits(EMISSIONS_H2()))


reset(hc)
run(hc)

# Format output 
fetchvars(hc, dates, vars) %>%
    mutate(year = year - IMPULSE_YR) %>%
    filter(year >= 0) %>%
    filter(year <= 100) %>%
    select(year, variable, value, units) ->
    h2_100_rslts
```


Integrate the pulse run results and calculate the AGWP100 

```{r}
# Calculate the GWP100 as the ratio of the cumulative RF change 100 years (integrated)
co2_integral <- summarise(co2_100_rslts, value = sum(value), .by = c("variable"))
h2_integral <- summarise(h2_100_rslts, value = sum(value), .by = c("variable"))

# This is consistent with the model mean GWP 100 due to CH4 from Sand et al.
# Supplementary Table 6
GWP100 <- round(h2_integral$value[h2_integral$variable == "RF_tot"] / co2_integral$value[co2_integral$variable == "RF_tot"], 3)


paste0("GWP100 H2: ", GWP100)
```

Looking at Sand et al. Supplementary Table 6 the GWP100 of hydrogen due to strat H2O ranges from 0.9 to 3.2 with a model mean of 2.1, which is not too far off what we are getting from Hector at the moment! 


RF response to the H2 pulse



```{r}
h2_100_rslts$type <- "H2"
co2_100_rslts$type <- "CO2"


h2_100_rslts %>%
    bind_rows(co2_100_rslts) %>%
    filter(variable == RF_TOTAL()) %>%
    ggplot(aes(year, value, color = type))+
    geom_line() +
    labs(title = "RF from 1Tg Emissions",
         subtitle = paste0("GWP100: ", GWP100),
         caption = "H2 impacts Tau OH",
         x = "years since emissions pulse",
         y = "Total RF (Wm-2)") +
    theme(legend.title = element_blank())
```
